---
title: "Project 1"
author: "Rachel Hardy"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: readable
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, show_col_types = FALSE)
```

# Data Processing

## First Steps

Step 0: Reading in the data.
```{r}
library(readr)
sheet1 <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv")
```
Step 1: Selecting certain columns.
```{r}
library(tidyverse)
newSheet1 <- sheet1 %>% select(Area_name, STCOU, ends_with("D")) %>% rename(area_name = Area_name)
newSheet1
```
Step 2: Converting data into long format.
```{r}
newSheet2 <- newSheet1 %>% pivot_longer(cols = ends_with("D"), names_to = "survey", values_to = "enrollment")
newSheet2
```
Step 3: Parse the string to pull out the year and convert the year into a numeric value. In addition, grab the first three characters and following four digits to create a new variable representing which measurement was grabbed. I will also be creating a variable for the type of value we have from the survey.
```{r}
newSheet2$year <- substr(newSheet2$survey, 8, 9) %>% as.Date(format = "%y") %>% format("%Y") %>% as.numeric()
newSheet2 <- newSheet2 %>% mutate(type = substr(survey, 4, 7), measurement = substr(survey, 1, 3))
newSheet3 <- newSheet2 %>% select("area_name", "STCOU", "measurement", "type", "year", "enrollment")
newSheet3
```
Step 4: Create two data sets. One that is only non-county level data and one that is only county level data.
```{r}
county <- newSheet3[c(grep(pattern = ", \\w\\w", newSheet3$area_name)), ]
class(county) <- c("county", class(county))
county

state <- newSheet3[-c(grep(pattern = ", \\w\\w", newSheet3$area_name)), ]
class(state) <- c("state", class(state))
state
```
Step 5: For the county level tibble, create a new variable that describes which state one of these county measurements corresponds to.
Note: I used information I found online to create the substrRight function in order to select the last two characters of a string.
```{r}
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x)) }

county <- county %>% mutate(state = substrRight(county$area_name, 2))
county
```
Step 6:  For the non-county level tibble, create a new variable called “division” corresponding to the state’s classification of division here. If row corresponds to a non-state (i.e. UNITED STATES), return ERROR for the division.
```{r}
state <- state %>% mutate(division = if_else(area_name %in% c("WASHINGTON", "IDAHO", "MONTANA", "OREGON", "WYOMING", "COLORADO", "UTAH", "NEVADA", "CALIFORNIA", "ALASKA", "HAWAII"), "West", 
  if_else(area_name %in% c("NORTH DAKOTA", "SOUTH DAKOTA", "NEBRASKA", "KANSAS", "MINNESOTA", "IOWA", "MISSOURI", "WISCONSIN", "ILLINOIS", "MICHIGAN", "INDIANA", "OHIO"), "Midwest", 
  if_else(area_name %in% c("ARIZONA", "NEW MEXICO", "TEXAS", "OKLAHOMA"), "Southwest",
  if_else(area_name %in% c("MARYLAND", "DELAWARE", "PENNSYLVANIA", "NEW JERSEY", "CONNECTICUT", "RHODE ISLAND", "NEW YORK", "MASSACHUSETTS", "NEW HAMPSHIRE", "VERMONT", "MAINE"), "Northeast",
  if_else(area_name %in% c("ARKANSAS", "LOUISIANA", "MISSISSIPPI", "ALABAMA", "FLORIDA", "GEORGIA", "SOUTH CAROLINA", "NORTH CAROLINA", "KENTUCKY", "TENNESSEE", "WEST VIRGINIA", "VIRGINIA", "DISTRICT OF COLUMBIA", "District of Columbia"), "Southeast",
          "ERROR"))))))
state
```

## Requirements

Step 1: Creating a function that does steps 0, 1, and 2 from the previous section. Make sure to add the optional argument!
```{r}
creatingLong <- function(url, var_name = "enrollment") {
  data1 <<- read_csv(url)
  data1 <<- data1 %>% select(Area_name, STCOU, ends_with("D")) %>% rename(area_name = Area_name)
  data1 <<- data1 %>% pivot_longer(cols = ends_with("D"), names_to = "survey", values_to = var_name)
  return(data1)
}

creatingLong("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
```
Step 2: Creating a function that takes in the output of step 2 and does step 3 from the previous section.
```{r}
parsingData <- function(url) {
  creatingLong(url)
  data1 <<- data1$year <- substr(data1$survey, 8, 9) %>% as.Date(format = "%y") %>% format("%Y") %>% as.numeric()
  data1 <<- data1$type <- substr(data1$survey, 4, 7)
  data1 <<- data1$measurement <- substr(data1$survey, 1, 3)
  data2 <<- data1 %>% select("area_name", "STCOU", "measurement", "type", "year", "enrollment")
  return(data2)
}

parsingData("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
```
Step 3: Creating a function to do step 5 from the previous section.
Note: This is the same function that I used in step 5 of the previous section!
```{r}
substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x)) }
```
Step 4: Creating a function to do step 6 from the previous section.
```{r}
creatingDivisions <- function(var1) {
  division = if_else(var1 %in% c("WASHINGTON", "IDAHO", "MONTANA", "OREGON", "WYOMING", "COLORADO", "UTAH", "NEVADA", "CALIFORNIA", "ALASKA", "HAWAII"), "West", 
  if_else(var1 %in% c("NORTH DAKOTA", "SOUTH DAKOTA", "NEBRASKA", "KANSAS", "MINNESOTA", "IOWA", "MISSOURI", "WISCONSIN", "ILLINOIS", "MICHIGAN", "INDIANA", "OHIO"), "Midwest", 
  if_else(var1 %in% c("ARIZONA", "NEW MEXICO", "TEXAS", "OKLAHOMA"), "Southwest",
  if_else(var1 %in% c("MARYLAND", "DELAWARE", "PENNSYLVANIA", "NEW JERSEY", "CONNECTICUT", "RHODE ISLAND", "NEW YORK", "MASSACHUSETTS", "NEW HAMPSHIRE", "VERMONT", "MAINE"), "Northeast",
  if_else(var1 %in% c("ARKANSAS", "LOUISIANA", "MISSISSIPPI", "ALABAMA", "FLORIDA", "GEORGIA", "SOUTH CAROLINA", "NORTH CAROLINA", "KENTUCKY", "TENNESSEE", "WEST VIRGINIA", "VIRGINIA", "DISTRICT OF COLUMBIA", "District of Columbia"), "Southeast",
          "ERROR")))))
}
```
Step 5: Creating a function that takes in the output from step 3 and creates the two tibbles in step 4, calls the above two functions (to perform steps 5 and 6), and returns two final tibbles.
```{r}
creatingCountyState <- function(url) {
  parsingData(url)
  
  county1 <<- data2[c(grep(pattern = ", \\w\\w", data2$area_name)), ]
  county1 <<- county1 %>% mutate(state = substrRight(county$area_name, 2))
  class(county1) <- c("county", class(county1))
  
  state1 <<- data2[-c(grep(pattern = ", \\w\\w", data2$area_name)), ]
  state1 <<- state1 %>% mutate(division = creatingDivisions(state1$area_name))
  class(state1) <- c("state", class(state1))
  
  return(list(county1, state1))
}

creatingCountyState("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
```
Step 6: Create the wrapper function!
```{r}
my_wrapper <- function(url, var_name = "enrollment") {
  creatingLong(url, var_name = var_name)
  parsingData(url)
  creatingCountyState(url)
  return(list(county1, state1))
}

my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
```

# Call It and Combine Your Data

Call the function you made two times to read in and parse the two .csv files mentioned so far. Be sure to
call the new value column the same in both function calls.
```{r}
EDU01 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv")
EDU02 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
```
Write a small function that takes in the results of two calls to your wrapper function and combines the tibbles
appropriately. There will be two final tibbles: one for all county level data and one for all state level data!
```{r}
combineTibbles <- function(list1, list2) {
  county01 <- as_tibble(list1[[1]])
  state01 <- as_tibble(list1[[2]])
  county02 <- as_tibble(list2[[1]])
  state02 <- as_tibble(list2[[2]])
  countyFinal <- dplyr::bind_rows(county01, county02)
  stateFinal <- dplyr::bind_rows(state01, state02)
  return(list(countyFinal, stateFinal))
}

EDU <- combineTibbles(EDU01, EDU02)
```

# Writing a Generic Function for Summarizing

Run the following code in the console!
```{r, eval = FALSE}
plot.function  #what is used for a class = function
getS3method("plot","data.frame")  #what is used for a class = data frame
```
For the state plotting function, I'll write a function that plots the mean value of the statistic (enrollment for this data set) across the years for each division. Observations with a division of "ERROR" will need to be removed.
```{r}
library(tidyverse)

stateFinalEDU <- as_tibble(EDU[[2]])

plot.state <- function(df, var_name = "enrollment") {
  df <- df %>% group_by(division, year) %>% summarize(mean = mean(get(var_name)))
  new_df <- df[!(df$division == "ERROR"),]
  
  ggplot(new_df, aes(x = year, y = mean, color = division)) + geom_line()
}
```
```{r, eval = FALSE, echo = FALSE}
plot.state(stateFinalEDU)
```
For the county plotting function, the function will allow the user to: specify the state of interest, giving a default value if not specified - determine whether the ‘top’ or ‘bottom’ most counties should be looked at with a default for ‘top’ - instruct how many of the ‘top’ or ‘bottom’ will be investigated with a default value of 5.
```{r}

countyFinalEDU <- as_tibble(EDU[[1]])

plot.county <- function(df, state_abbrev = "AL", var_name = "enrollment", top = TRUE, bottom = FALSE, x = 5) {
  df1 <<- df %>% filter(state == state_abbrev)
  df2 <<- df1 %>% group_by(area_name) %>% mutate(average = mean(get(var_name)))
  df3 <<- if (top) {
            df2[order(df2$average, decreasing = TRUE),]
            df2[1:x, ]
          }
          if (bottom) {
            df2[order(df2$average, decreasing = FALSE),]
            df2[1:x, ]
          }
  
  ggplot(df3, aes(x = year, y = get(var_name), color = area_name)) + geom_line()
}
#For the life of me I couldn't figure this out completely! :(
```
```{r, eval = FALSE, echo = FALSE}
plot.county(countyFinalEDU, state_abbrev = "NC", top = TRUE, x = 40)
```

# Put It Together!

## EDU Data Sets

Putting everything together to create the the EDU data sets.
```{r}
EDU01 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv")
EDU02 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")

EDU <- combineTibbles(EDU01, EDU02)

stateFinalEDU <- as_tibble(EDU[[2]])
countyFinalEDU <- as_tibble(EDU[[1]])
stateFinalEDU
countyFinalEDU
```
Using the plot functions on both the state and county final data sets.
```{r}
plot.state(stateFinalEDU)
plot.county(countyFinalEDU, state_abbrev = "PA", x = 7) #PA, top, 7 observations
plot.county(countyFinalEDU, state_abbrev = "PA", bottom = TRUE, x = 4) #PA, bottom, 4 observations
plot.county(countyFinalEDU) #Using all the default values
plot.county(countyFinalEDU, state_abbrev = "MN", x = 10) #MN, top, 10 observations
```

## PST Data Sets

Run the data processing (wrapper) function on the four urls given.
```{r}
PST01 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01a.csv")
PST02 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01b.csv")
PST03 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01c.csv")
PST04 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01d.csv")
```
Run the data combining function to put these into one object (with two data frames).
```{r}
PST0102 <- combineTibbles(PST01, PST02)
PST0304 <- combineTibbles(PST03, PST04)
PST <- combineTibbles(PST0102, PST0304)

stateFinalPST <- as_tibble(PST[[2]])
countyFinalPST <- as_tibble(PST[[1]])
stateFinalPST
countyFinalPST
```
Use the plot function on the PST state data frame.
```{r}
plot.state(stateFinalPST)
```

Use the plot function on the PST county data frame.
```{r}
plot.county(countyFinalPST, state_abbrev = "CT", x = 6) #CT, top, 6 observations
plot.county(countyFinalPST, state_abbrev = "NC", bottom = TRUE, x = 10) #NC, bottom, 10 observations
plot.county(countyFinalPST) #Using all the default values
plot.county(countyFinalPST, state_abbrev = "MN", x = 4) #MN, top, 4 observations
```


